Meeting Notes for Week 1/23/20 - 1/30/20

### 2020-01-30 notes/weekly_notes/1.30.20.MeetingNotes.tex
added higher-order testing rate overly compactly supported Sobolev functions

\begin{document}
\title{Graph Testing}
\author{Alden Green}
\date{\today}
\maketitle

This document details the current status of the graph testing project. We divide by section based on the testing problem under consideration. At the end of each section, we list some areas we are interested in investigating. All proofs are left until the end.

We establish some notation that we will use throughout. A graph $G = (V,W)$ consists of a collection of vertices $V = \{v_1,\ldots,v_n\}$, and a weight matrix $W \in \Reals^{\abs{V} \times \abs{V}}$ with $ij$th entry $W_{ij}$ encoding affinity between $v_i$ and $v_j$. We will sometimes write $W_G$ to make it clear which graph a weight matrix is associated with. When the weight matrix $W_G$ consists only of $0$s and $1$s, we will use the equivalent representation $G = (V,E(G))$ when convenient; here $E(G) = V \times V \cap \{(v_i,v_j): W_{ij} = 1\}$.

We will focus our attention on neighborhood graphs, and operators associated with them. For a kernel $K:\Reals \to \Reals$ and radius $r > 0$, we define $G_{n,K}=(V,W)$ to be the neighborhood graph with vertices $V = X$, and weights $W_{ij} = K(\norm{x_i - x_j}/r)$. Of particular interest to us will be the \emph{random geometric graph} $G_{n,r}$, the neighborhood graph constructed using the uniform kernel $K(z) := \1\{z \leq 1\}$. The weight matrix $W(G_{n,r})$ consists only of $0$s and $1$s, and we will thus use the equivalent representation $G_{n,r} = (X,E(G_{n,r}))$.

The primary operator we are interested in (at this point) is the graph Laplacian. Let $D_G$ be the diagonal degree
matrix of a graph $G$, with entries $D_{uu} := \sum_{v \in V} W_{uv}$. The graph Laplacian is then $L_G = D_G - W_G$. Let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $k$th eigenvalue $\lambda_k(G)$.

\section{Regression goodness-of-fit testing.}

Suppose we observe samples $(y_i,x_i)$ for $i = 1,\ldots,n$, where conditional on $X$ the responses $Y = \{y_1,\ldots,y_n\}$ are assumed to follow the model
\begin{equation}
\label{eqn:regression_known_variance}
y_i = f(x_i) + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
\end{equation} 
and our task is to distinguish
\begin{equation*}
\mathbf{H}_0: f = f_0 := 0 \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: f \neq f_0
\end{equation*}

We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
\begin{equation*}
\mathcal{R}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, f \neq f_0} \Ebb_f(1 - \phi).
\end{equation*}
The worst-case risk may be quite close to $1$ unless we enforce some separation between null and alternative spaces. A more realistic measure of performance is therefore
\begin{equation*}
\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_{\Leb^2} \geq \epsilon} \Ebb_f(1 - \phi).
\end{equation*}

\subsection{Test Statistics.}
We list the test statistics we use for the regression testing problem.

\paragraph{Eigenvector projection test statistic.}
The graph Laplacian eigenvector projection test is a truncated-series test. Suppose we are given a graph $G = (X,W)$ defined on the sample points $X$.  Letting $\kappa$ be some integer between $1$ and $n$, our test projects $Y$ onto the eigenvectors $v_1(G),\ldots,v_{\kappa}(G)$, and then takes the empirical norm of this projection to be the test statistic. Formally,
\begin{equation*}
T_{\mathrm{spec}}(G) := \frac{1}{n} \sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} y_i v_{k,i}(G)\Biggr)^2
\end{equation*}
Then the test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau$.

\subsection{Random design.}
In the random design case, we assume that each design point $x_i$ is sampled independently from some distribution $P$. Formally, our model for the samples is the following:
\begin{equation}
\label{eqn:regression_random_design_known_variance}
x_1,\ldots,x_n \overset{\textrm{i.i.d}}{\sim} P,~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)~\textrm{and}~ \varepsilon \perp X,~~ y_i = f(x_i) + \varepsilon_i.
\end{equation}

In Theorem~\ref{thm:sobolev_testing_rate_order1} we show that under some typical regularity conditions on $P$, the neighborhood graph Laplacian eigenvector projection test $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ for $d = 1,2,3$ and $\Xset = [0,1]^d$. Let $p_d$ equal $3/4$ when $d = 2$ and equal $1/d$ for $d \geq 3$.

\begin{theorem}
	\label{thm:sobolev_testing_rate_order1}
	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}. Let $R > 0$ and $b \geq 1$ be fixed constants, and $d = 1,2$ or $3$. Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x)$ is bounded away from zero and infinity, 
	\begin{equation*}
	0 < p_{\min} < p(x) < p_{\max} < \infty,~~ \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	and the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choices
	\begin{equation*}
	c \frac{(\log n)^{p_d}}{n^{1/d}} \leq r(n) \leq n^{-4/((4 + d)(2+d))}, ~\kappa = (nR)^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	for $c$ a constant that depends only on $\Xset, p_{\min}$ and $p_{\max}$. Then the following statements holds for every $n$ sufficiently large: there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
	\begin{equation}
	\label{eqn:sobolev_testing_rate_order1}
	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4 + d)}\} \cdot n^{-4/(4 + d)}
	\end{equation}
	the worst-case risk is upper bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_order1_1}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); W^{1,2}(\mathcal{X};R)) \leq \frac{c_2}{b}.
	\end{equation}
\end{theorem}

The testing problem in the random design case with known variance fundamentally changes in the low-smoothness regime, where the number of bounded derivatives is small compared to the dimension $d$. In the case where $f \in W^{1,2}(\Xset;R)$ is assumed to have one derivative which is bounded in $\Leb^2(\Xset)$ norm, this transition occurs at $d = 4$. We shall discuss testing in the low-smoothness regime shortly.

Theorem~\ref{thm:sobolev_testing_rate_order1} establishes that a truncated series test using graph eigenvectors performs just as well, in a minimax sense, as a truncated series test using a classic Fourier series; for analysis of the latter test as well as derivation of the minimax rate of nonparametric goodness-of-fit testing over Sobolev classes see \textcolor{red}{(Ingster)}. Intuitively, we can understand this by thinking of the graph eigenvectors as estimates of eigenfunctions of a density-weighted continuum Laplacian operator. When the density is uniform, and the domain $\Xset$ is the unit cube, under appropriate boundary conditions these eigenfunctions are themselves elements of the Fourier basis. When the density is merely close to uniform, the eigenfunctions no longer belong to the Fourier basis; nevertheless, they share enough properties with the latter that the resulting minimax rate is unchanged. Theorem~\ref{thm:sobolev_testing_rate_order1} shows that the additional error due to using \textit{estimates} of these eigenfunctions does not fundamentally change the minimax rate.

When we additionally believe $W^{s,2}(\Xset)$, we know (again from Ingster) that the minimax testing rate is $n^{-4s/(4s + d)}$. Now, however, our estimates of the eigenvectors no longer track the derivatives finely enough at the boundaries of $\Xset$ to establish the same minimax rate over all $W^{s,2}(\mathcal{X})$. However, if we assume $f$ and its derivatives are supported only on the interior of $\mathcal{X}$, we recover the usual rate. Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev balls $W_{\sigma}^{s,2}(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.

\begin{theorem}
	\label{thm:sobolev_testing_rate}
	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}. Let $R,\sigma > 0$, $b,s,d \geq 1$ be fixed constants, with $s$ and $d$ integers. Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e. there exists $p_{\min} > 0$ such that 
	\begin{equation*}
	p_{\min} < p(x),~~ \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	Then the following statement holds for all $n$ sufficiently large: if the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choice
	\begin{equation*}
	n^{-1/(2(s-1) + d)} \leq r(n) \leq \min\bigl\{n^{-4/((4s + d)(2+d))},\sigma/s\bigr\}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	then there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
	\begin{equation}
	\label{eqn:sobolev_testing_rate}
	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-{4s}/(4s + d)}
	\end{equation}
	the worst-case risk is upper bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_1}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W_{\sigma}^{s,2}(\mathcal{X};R)) \leq \frac{c_2}{b}.
	\end{equation}
\end{theorem}

The restriction that $p$ be supported on the unit cube is mostly for convenience. If instead $p$ is supported on any compact set with Lipschitz boundary, the Theorem statement should hold (up to constants), with only a few modifications to our proofs. However, we do not work through the details.

The restriction $4s > d$ is a fundamental consequence of the tightness of the Sobolev embedding theorem. When $4s \geq d$ the compact embedding
\begin{equation*}
W_d^{s,2}(\mathcal{X}) \subseteq \mathcal{L}_d^4(\mathcal{X}) 
\end{equation*}
does not hold (as it does when $d < 4s$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
\begin{proposition}
	\label{prop:L4_testing_rate}
	If $f \in \Leb^4(\mathcal{X};R)$, there exists a constant $c$ such that if
	\begin{equation}
	\label{eqn:L4_testing_rate}
	\epsilon^2 > b \cdot n^{-1/2}
	\end{equation}
	then the test
	\begin{equation*}
	\phi_{\mathrm{mean}} = \1\{\frac{1}{n}\sum_{i = 1}^{n} y_i^2 \geq 1\}
	\end{equation*}
	has worst-case risk
	\begin{equation*}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}. 
	\end{equation*}
\end{proposition}
Note that when $4s > d$ the critical radius~\eqref{eqn:sobolev_testing_rate} is smaller than \eqref{eqn:L4_testing_rate}.

\subsection{Analysis}

To prove Theorem~\ref{thm:sobolev_testing_rate}, we show that there exists a high-probability set $E \subseteq \Xset^n$ such that conditional on $X \in E$, the test $\phi_{\spec}(G_{n,r})$ has small risk. Since $G_{n,r}$ is a function only of $X$ and not of $Y$, this amounts to reasoning about the behavior of the test $\phi_{\spec}$ over a fixed graph $G = (X,E)$, where we observe
\begin{equation}
y_i = \beta_i + \varepsilon_i,~~\varepsilon_i \sim \mathcal{N}(0,1)
\end{equation}
for some fixed $\beta \in \Reals^n$.  In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}(G)$. Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$.

\subsection{Areas to Investigate}

\begin{enumerate}[(i)]
	\item When $d \geq 4$, is it true that there does not exist a uniformly consistent
	test over the Sobolev ball $W_d^{1,2}(\mathcal{X};R)$?
	\item Assume the distribution $P$ is supported on a manifold $\mathcal{M}$ of intrinsic dimension $s < d$. Does $\mathrm{\phi_{\mathrm{spec}}}$ display adaptivity to the intrinsic dimension of $\mathcal{M}$?
	\item Assume that $f$ belongs to the Holder space $C_d^s(\mathcal{X})$. Moreover, suppose that instead of observing ${y_i}$ according to the regression testing model \eqref{eqn:regression_known_variance}, we observe
	\begin{equation*}
	y_i = f(x_i) + \sigma \varepsilon_i, ~\varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
	\end{equation*}
	where $\sigma > 0$ is unknown. When $d \geq 4$, what are the minimax regression testing rates over $C_d^1(\mathcal{X};R)$? Is the test $\phi_{\mathrm{spec}}$ minimax optimal, when the tuning parameters $r$ and $\kappa$ are appropriately chosen?
\end{enumerate}

\section{Two-sample density testing.}
In the two-sample density testing problem, we observe independent samples $Z = z_1,\ldots,z_N \sim P$ and $Y = y_1,\ldots,y_M \sim Q$, where $P$ and $Q$ are distributions over $\Reals^d$ with densities $p$ and $q$, respectively, and $N \sim \textrm{Bin}(n,1/2)$. Our goal is to distinguish the hypotheses
\begin{equation*}
\mathbf{H}_0: P = Q \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: P \neq Q
\end{equation*}
and we again evaluate our performance using worst-case risk; letting $\phi:\Reals^{N + M} \to \{0,1\}$, 
\begin{equation*}
\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \inf_{p \in \mathcal{H}}\Ebb_{p,p}(\phi) + \sup_{\substack{p,q \in \mathcal{H} \\ \norm{p - q}_{\Leb^2} \geq \epsilon}} \Ebb_{p,q}(1 - \phi).
\end{equation*}

\subsection{Test statistics.}

We suggest several two-sample test statistics.

\paragraph{Eigenvector projection test statistic.}

It is straightforward to adapt the test statistic $T_{\mathrm{spec}}$ to the two-sample testing problem. Concatenate the samples $Z$ and $Y$ in $X = (z_1,\ldots,z_N,y_1,\ldots,y_M)$, and define $T_{\mathrm{spec}}^{(2)}$ to be
\begin{equation}
\label{eqn:graph_spectral_projections_2}
T_{\mathrm{spec}}^{(2)} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2, ~~\textrm{where}~~ a = (\underbrace{N^{-1},\ldots,{N^{-1}}}_{\textrm{length } N},\underbrace{-M^{-1},\ldots,-M^{-1}}_{\textrm{length } M})
\end{equation}

For convenience, we state our following two test statistics with respect to the empirical norm $\norm{\theta}_n = n^{-1/2}\norm{\theta}_2$ for $\theta \in \Reals^n$. They will each depend on a tuning parameter $\lambda > 0$.
\paragraph{Graph Sobolev IPM.}
Letting $C_n := nr^{(d + 2)/2}$ and
\begin{equation*}
\Theta_{1,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B\theta}_2 \leq 1\} 
\end{equation*}
we define the \emph{graph Sobolev IPM} to be
\begin{equation}
\label{eqn:sobolev_IPM}
T_{\textrm{sob}} := \sup_{\substack{\theta \in \Theta_{1,2} \\ \lambda \norm{\theta}_n \leq 1}} \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}. 
\end{equation}

\paragraph{Graph Total Variation IPM.}
Letting $C_n' := n^{2}r^{(d + 1)}$ and 
\begin{equation*}
\Theta_{1,1} := \{\theta \in \Reals^n:~ (C_n')^{-1} \norm{B\theta}_1 \leq 1\}
\end{equation*}
we define the \emph{graph Total Variation} IPM to be
\begin{equation}
\label{eqn:total_variation_IPM}
T_{\mathrm{TV}} := \sup_{\substack{\theta \in \Theta_{1,1}, \\ \lambda \norm{\theta}_n \leq 1} } \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}, \quad
\end{equation}

\subsection{Current results.}

In Theorem~\ref{thm:twosample_sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}}^{(2)} := \1\{T_{\mathrm{spec}}^{(2)} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ when $d = 1$.

\begin{theorem}
	\label{thm:twosample_sobolev_testing_rate}
	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d = 1$.  Suppose that $\mu = (P + Q)/2$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]$ with density functions $\rho$ bounded above and below by constants, i.e
	\begin{equation*}
	0 < \rho_{\min} < \rho(x) < \rho_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	Then the following statement holds: if the test $\phi_{\spec}^{(2)}$ is performed with parameter choices 
	\begin{equation*}
	r = \log^a n \cdot \left(\frac{\log n}{n}\right), ~\kappa = n^{2/5}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	then there exists a constant $c$ which may depend on $R,p_{\max},q_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
	\begin{equation}
	\label{eqn:twosample_sobolev_testing_rate}
	\epsilon^2 \geq c \cdot b \cdot n^{-4/5} (\log n)^{h(a,1)}
	\end{equation}
	the worst-case risk is upper bounded
	\begin{equation}
	\label{eqn:twosample_sobolev_testing_rate_1}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
	\end{equation}
\end{theorem}

We prove Theorem~\ref{thm:twosample_sobolev_testing_rate} by relating the density testing problem to a regression testing problem with a certain type of structured noise, and then proceeding along similar lines to the proof of Theorem~\ref{thm:sobolev_testing_rate}. To pursue this strategy, we require the eigenvectors to satisfy a certain type of incoherence condition; this is in constrast to the regression testing problem with known variance, where we did not require the eigenvectors to be smooth in any sense.

\subsection{Areas to Investigate}

\begin{enumerate}[(i)]
	\item When $1 < d < 4$, is the test $\phi_{\spec}^{(2)}$ minimax optimal?
	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $C_d^1(\mathcal{X};R)$ for all values of $d$?
	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $W_d^{1,2}(\mathcal{X};R)$ when $d \leq 4$?
	\item Is the test statistic \eqref{eqn:graph_spectral_projections_2}, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$? 
	\item Modify the test statistic \eqref{eqn:sobolev_IPM} by replacing the function class $\Theta_{1,2}$ with
	\begin{equation}
	\Theta_{s,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B^{(s)}\theta}_2 \leq 1\} 
	\end{equation}
	Is the modified test statistic, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
	\item What is the minimax testing rate over $BV_d^{1}(\mathcal{X};R)$? Does it exhibit a phase transition analogous to the minimax estimation rate over bounded variation spaces?
	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over $BV_d^{1}(\mathcal{X};R)$?
	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over Sobolev and Holder function classes?
\end{enumerate}

\section{Definitions and Notation}

Here we collect definitions of some common function spaces and graph operators.

\subsection{Function Spaces}

\paragraph{Lebesgue spaces.}

We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if 
$$\norm{f}_{\mathcal{L}^p(\mathcal{X})} := \int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$$
and we let 
\begin{equation*}
\mathcal{L}^p(\mathcal{X};R) = \set{f \in \mathcal{L}^p(\mathcal{X}): \norm{f}_{\mathcal{L}^p(\mathcal{X})} < R}
\end{equation*}
be a ball in the Lebesgue space.

\paragraph{Holder spaces.}

For a given $s > 0$, let $\ell = \floor{s}$ be the largest integer strictly less than $s$. Then the $s$th Holder norm is given by
\begin{equation*}
\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} < s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = \ell} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2^{s - \ell}}
\end{equation*}
and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $\ell$ times continuously differentiable with finite Holder norm $\norm{\cdot}_{C_d^{s}(\mathcal{X})}$. Denote the Holder ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C_d^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.

\paragraph{Sobolev spaces.}

For a given $s > 0$, the Sobolev space $W^{s,2}(\mathcal{X})$ consists of all functions $f \in \Leb^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$. The Sobolev $\{s,2\}$ norm is then 
\begin{equation*}
\norm{f}_{W^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
\end{equation*}
and for a given $R > 0$, the corresponding ball is $W^{s,2}(\Xset; R) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq R}$.

For a given $\sigma > 0$, let $\Xset_{\sigma} := \set{x \in \Xset: \mathrm{dist}(x,\partial\Xset) > \sigma}$. Let $W_{\sigma}^{s,2}(\Xset;L)$ consist of all those functions $f \in W^{s,2}(\Xset)$ such that $\mathrm{supp}(f) \subseteq \Xset_{\sigma}$.

\paragraph{Bounded Variation spaces.}

For a function $f \in L^1(\mathcal{X})$ the \emph{total variation} semi-norm of $f$ is
\begin{equation*}
TV(f;\mathcal{X}) := \sup \left\{ \int_{\mathcal{X}} f \, \Xsetive \, \psi \,dx : \psi \in C_c^1(\mathcal{X}; \Reals^d), \abs{\psi} \leq 1 \right\};
\end{equation*}
and we write $BV_d(\mathcal{X})$ for the subset of functions $f \in L^1(\mathcal{X})$ which have bounded norm
\begin{equation*}
\norm{f}_{BV_d(\mathcal{X})} := \norm{f}_{\infty} + TV(f;\mathcal{X}).
\end{equation*}
For a given $R > 0$, the corresponding ball is $BV_d^{1}(\mathcal{X};R) = \set{f: \norm{f}_{BV_d(\mathcal{X})} \leq R}$.

\subsection{Graph Operators}
Let $s \geq 1$ be an integer. The $s$th-order difference operator on $G_{n,r}$, denoted $B^{(s)}$, is defined by
\begin{equation*}
B^{(s)} :=
\begin{cases}
L^{s/2},& ~~ s \textrm{ even} \\
BL^{(s - 1)/2},& ~~ s \textrm{ odd.}
\end{cases}
\end{equation*}

\subsection{Kernels}
We say a kernel function $K(\cdot)$ is a $2$nd order kernel if $K$ is compactly supported on $B(0,1)$, uniformly upper bound $\abs{K(x)} \leq K_{\max} < \infty$ for all $x \in B(0,1)$, and
\begin{equation*}
\int K(x) \,dx = \nu_d,~~ \int x K(x) \,dx = 0.
\end{equation*}
Note that the uniform kernel, defined as $K(x) = \1\{\norm{x} \leq 1\}$, is a $2$nd order kernel.

\subsection{Notation}

We will treat $f$ interchangeably as a function $f:\Rd \to \Reals$, and as a vector $f = (f(x_1),\ldots,f(x_n))$. It should always be clear from context how we are using $f$.
For sequences $(a_n), (b_n)$, we say $a_n = O(b_n)$ if there exists a constant $c$ such that $a_n \leq cb_n$ for all $c$. We use the notation $a \wedge b$ for the minimum of $a$ and $b$.

We use the notation $\Lambda(H)$ to denote the spectrum of a matrix $H$, and $\lambda_k(H)$ to denote the $k$th smallest eigenvalue of $H$.

@@

\subsection{Proof of Lemma~\ref{lem:fixed_graph_testing}}

In this proof, we will drop all notational dependence on the graph $G$ for ease of reading.

To prove Lemma~\ref{lem:fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.

>>>>

\begin{document}
\title{Meeting Notes for Week 1/23/20 - 1/30/20}
\author{Alden Green}
\date{\today}
\maketitle

Suppose we observe samples $(y_i,x_i)$ for $i = 1,\ldots,n$, where $x_1,\ldots,x_n$ are sampled independently from a distribution $P$ with density $p$, and conditional on $X$ the responses $Y = \{y_1,\ldots,y_n\}$ are assumed to follow the model
\begin{equation}
\label{eqn:regression_random_design_known_variance}
y_i = f(x_i) + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
\end{equation} 
Our task is to distinguish
\begin{equation*}
\mathbf{H}_0: f = f_0 := 0 \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: f \neq f_0
\end{equation*}
and we worst-case risk to assess performance: for a test $\phi$ and function class $\mathcal{H}$,
\begin{equation*}
\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_{\Leb^2} \geq \epsilon} \Ebb_f(1 - \phi).
\end{equation*}

Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that the graph spectral test 
\begin{equation*}
\phi_{\spec}(G_{n,r}) := \1\bigl\{T_{\textrm{spec}}(G_{n,r}) \geq \tau\bigr\},~~ T_{\textrm{spec}}(G_{n,r}) := \frac{1}{n}\sum_{k = 1}^{\kappa} \biggl(\sum_{i = 1}^{n} v_{k,i}(G_{n,r}) y_i\biggr)
\end{equation*} is a minimax optimal test over the Sobolev balls $W_{\sigma}^{s,2}(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.

Here, $W_{\sigma}^{s,2}(\Xset;L)$ consists of all those functions $f \in W^{s,2}(\Xset)$ such that $\mathrm{supp}(f^{(\alpha)}) \subseteq \Xset_{\sigma}$ for each $\abs{\alpha} \leq s$.

\begin{theorem}
	\label{thm:sobolev_testing_rate}
	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}. Let $R,\sigma > 0$, $b,s,d \geq 1$ be fixed constants, with $s$ and $d$ integers. Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e. there exists $p_{\min} > 0$ such that 
	\begin{equation*}
	p_{\min} < p(x),~~ \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	Then the following statement holds for all $n$ sufficiently large: if the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choices
	\begin{equation*}
	n^{-1/(2(s-1) + d)} \leq r(n) \leq \min\bigl\{n^{-4/((4s + d)(2+d))},\sigma/s\bigr\}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	then there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
	\begin{equation}
	\label{eqn:sobolev_testing_rate}
	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-{4s}/(4s + d)}
	\end{equation}
	the worst-case risk is upper bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_1}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W_{\sigma}^{s,2}(\mathcal{X};R)) \leq \frac{c_2}{b}.
	\end{equation}
\end{theorem}

@@

\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate}}

We want to apply Lemma~\ref{lem:fixed_graph_testing}, which analyzes testing over an arbitrary graph $G$, to the case where $G = G_{n,r}$. In order to do this, we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$. In particular, we will show that the following statements each hold with probability at least $1 - c/b$ for sufficiently large $n$ (Here and in what follows, $c$ denotes a constant which is fixed in $n$ and $f$ and does not depend on $b$, but may depend on other fixed quantities such as $d$, $s$, etc.): 
\begin{enumerate}[label=(E\arabic*)]
	\item 
	\label{event:discrete_sobolev_norm}
	\textbf{Graph Sobolev norm:} For any $n^{-1/(2(s - 1) + d)}\leq r \leq \sigma/s$,
	\begin{equation}
	\label{eqn:continuous_to_discrete_sobolev_norm}
	f^T L^s f \leq c \cdot b \cdot \norm{f}_{W^{s,2}(\Xset)}^2 \cdot n^{s + 1} r^{s(d + 2)} 
	\end{equation}
	\item 
	\label{event:eigenvalue_tail_decay_2}
	\textbf{Eigenvalue tail bound:} There exists a constant $c$ such that for any $\kappa = 1,\ldots,n$, if $\max\{ 3 d_{\infty}(X,\ol{X}),n^{-1/d}\} \leq r \leq \kappa^{-2/(d(2 +d))}$ then,
	\begin{equation}
	\label{eqn:eigenvalue_tail_bound_2}
	\lambda_{\kappa} \geq c \cdot n r^{d + 2} \kappa^{2/d}
	\end{equation}
	In particular, the inequality~\eqref{eqn:eigenvalue_tail_bound_2} is satisfied for sufficiently large $n$ when $\kappa = n^{2d/(4s + d)}$ and $c(\log n/n)^{1/d} \leq r \leq n^{-4/((2+d)(4s + d))}$. 
	\item 
	\label{event:l2_norm}
	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-2s/(4s + d)}$ and $4s > d$,
	\begin{equation}
	\label{eqn:l2_to_empirical_norm}
	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
	\end{equation}
\end{enumerate}
